Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
ggplot(data = subset(pf, !is.na(gender)), aes(x = gender, y = age)) +
  geom_boxplot() + 
  stat_summary(fun.y = mean, geom = "point", shape = 4)

ggplot(data = subset(pf, !is.na(gender)), aes(x = age, y = friend_count)) +
  geom_line(aes(color = gender), stat = "summary", fun.y = median)

library(dplyr)

pf <- na.omit(pf)
age_groups <- group_by(pf, age, gender)
pf.fc_by_age_gender <- summarise(age_groups,
          mean_friend_count = mean(friend_count),
          median_friend_count = median(friend_count),
          n = n())
head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(pf.fc_by_age_gender, aes(y = median_friend_count, x = age)) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}

library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(pf.fc_by_age_gender.wide, aes(y = female/male, x = age)) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined = ceiling(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)
pf$year_joined.bucket = cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
library(ggplot2)
ggplot(data = subset(pf, !is.na(year_joined.bucket)), aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```

***

### Friending Rate
Notes:

```{r Friending Rate}
aux <- subset(pf, tenure >= 1)
rate <- aux$friend_count/aux$tenure
median(rate)
max(rate)
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
length(aux$friendships_initiated/aux$tenure)

ggplot(data = aux, aes(x = tenure, y = (friendships_initiated/tenure))) + 
  geom_line(aes(color = year_joined.bucket), fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)
yo$id <- factor(yo$id)
str(yo)

ggplot(data = yo, aes(x = price)) + 
  geom_histogram()
qplot(data = yo, x = price, fill = I('#F79420'))
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
qplot(x = all.purchases, data = yo, binwidth = 1, fill = I('#099DD9'))

ggplot(aes(x = time, y = price), data = yo) + geom_point()
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(400)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), 
       data = subset(yo, id %in% sample.ids)) + 
  facet_wrap( ~ id) + 
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

